Carrier phase and amplitude estimation for phase shift keying using pilots and data

### 2012-08-18 paper1.tex
added to papers
started filling in the simulation section.

%opening
\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data}
\author{Robby McKilliam, Andre Pollok, Bill Cowley  
\thanks{
Supported under the Australian Government’s Australian Space Research Program.
Robby McKilliam, Andre Pollok and Bill Cowley are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
}}

@@

For Theorem~\ref{thm:normality} the integral defining $g(\phi)$ must exist at $\phi = \tfrac{2\pi}{M}k + \frac{\pi}{M}$ for $k = 0, \dots, M-1$ in order the constant $H$ to exist.  The next proposition shows that the assumptions made in Theorem~\ref{thm:normality} are strong enough to ensure that the integral does exist at these points.  At the same time we derive some properties that will be useful during the proof of Theorem~\ref{thm:normality}.

@@

It remains to prove part 2 of the theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
\begin{equation}\label{eq:rhoGLZ}
\hat{\rho} = \frac{1}{L}\Re(Z(\hat{\theta})) = \rho_0 G_L(\hat{\lambda}_L).
\end{equation}  
Lemma~\ref{lem:GLtoG0} shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.

@@

We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Let $T_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
\begin{align*}
\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) + T_L(\hat{\lambda}_L) - G(0) \big).
\end{align*}
Lemma~\ref{lem:empiricprocforrho} shows that 
\[
\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L,
\]
where $X_L = \sqrt{L}\big( G_L(0) - T_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
\[
%\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1).
\]
It follows that
\[
%\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
\sqrt{L} m_L =  \rho_0 X_L + o_P(1). 
\]
Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$ where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.  Thus, the distribution of $\sqrt{L} m_L$ converges to the normal with zero mean and variance $\rho_0^2(p B_1 + d B_2)$ as required.  Because $X_L$ does not depend on $\hat{\lambda}_L$, it follows that $\covar(X_L, \sqrt{L}\hat{\lambda}_L) = 0$, and so,
\[
\covar(\sqrt{L}\hat{m}_L, \sqrt{L}\hat{\lambda}_L) = \covar(\rho_0 X_L + o_P(1), \sqrt{L}\hat{\lambda}_L) \rightarrow 0
\]
as $L \rightarrow \infty$.  It remains to prove the lemmas that we have used.  These lemmas are proved in~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.

@@

Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.  The joint density function of the real and imaginary parts is then
\[
p(x,y) = \frac{1}{2\pi\sigma^2}e^{-(x^2 + y^2)/\sigma^2}
\]
Theorem~\ref{thm:consistency}~and~\ref{thm:normality} hold and, since the distribution of $w_1$ is circularly symetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
%and the joint density function of the real and imaginary parts of $Z$ is
%\[
%p_Z(x,y) = \rho_0^2 p( \rho_0 (x - 1), \rho_0 y ) = \frac{\kappa^2}{2\pi}e^{-\kappa^2(x ^2 - 2x + 1 + y^2) }
%\] 
%where $\kappa = \tfrac{\rho_0}{\sigma}$.  
It can be shown that
\[
g(\phi) = \frac{\cos^2(\phi)}{4\sqrt{\pi}}\left( \frac{e^{-\kappa^2} }{\sqrt{\pi}} + \kappa \Psi(\phi)  e^{-\kappa^2\sin^2(\phi)}\cos(\phi) \right)
\]
where $\kappa = \rho_0/\sigma$ and
\[
\Psi(\phi) = 1 + \erf(\kappa \cos(\phi)) = 1 + \frac{2}{\sqrt{\pi}}\int_0^{\kappa \cos(\phi)} e^{-t^2} dt .
\]
The value of $A_1, A_2, B_1$ and $B_2$ can be efficiently computed by numerical integration using these formula.

@@

%\subsection{The coherent setting}

Our first set of simulations considers the case where there are a number of pilot symbols.

@@

%\subsection{The noncoherent setting}

%BPSK plots
% \begin{figure}[tp]
% 	\centering
% 		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
% 		\caption{Amplitude error for BPSK}
% 		\label{fig:plotamp}
 %\end{figure}

\begin{figure}[tp]
	\centering
		\includegraphics[width=\linewidth]{code/data/plotncM2-2.mps}
		\caption{Phase error for BPSK}
		\label{fig:plotphase}
\end{figure}

% \begin{figure}[tp]
% 	\centering
%         \includegraphics{code/data/plotM2-3.mps}
% 		\caption{complex amplitude error for BPSK}
% 		\label{fig:plotcamp}
% \end{figure}

%QPSK plots
 %\begin{figure}[tp]
% 	\centering
% 		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
% 		\caption{Amplitude error for QPSK}
% 		\label{fig:plotamp}
% \end{figure}

\begin{figure}[tp]
	\centering
		\includegraphics[width=\linewidth]{code/data/plotncM4-2.mps}
		\caption{Phase error for QPSK}
		\label{fig:plotphase}
\end{figure}

%8PSK plots
% \begin{figure}[tp]
% 	\centering
% 		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
% 		\caption{Amplitude error for 8PSK}
% 		\label{fig:plotamp}
% \end{figure}

\begin{figure}[tp]
	\centering
		\includegraphics[width=\linewidth]{code/data/plotncM8-2.mps}
		\caption{Phase error for 8PSK}
		\label{fig:plotphase}
\end{figure}

>>>>

%opening
\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data}
\author{Robby McKilliam, Andre Pollok, Bill Cowley  
\thanks{
Supported under the Australian Government’s Australian Space Research Program.
Robby McKilliam, Andre Pollok and Bill Cowley are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.  Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.  Barry~Quinn is with the Department of Statistics, Macquarie University, Sydney, NSW, 2109, Australia.
}}

@@

For Theorem~\ref{thm:normality} the integral defining $g(\phi)$ must exist at $\phi = \tfrac{2\pi}{M}k + \frac{\pi}{M}$ for $k = 0, \dots, M-1$ so that the constant $H$ exists.  The next proposition shows that the assumptions made in Theorem~\ref{thm:normality} are strong enough to ensure that the integral does exist at these points.  At the same time we derive some properties that will be useful during the proof of Theorem~\ref{thm:normality}.

@@

It remains to prove part 2 of the theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
\begin{equation}\label{eq:rhoGLZ}
\hat{\rho} = \frac{1}{L}\Re(Z(\hat{\theta})) = \rho_0 G_L(\hat{\lambda}_L).
\end{equation}  
Lemma~\ref{lem:GLtoG0} shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.  It remains to prove Lemmas~\ref{lem:convtoexpGlamL} and~\ref{lem:GLtoG0}.  These are proved in Section~\ref{sec:proof-almost-sureappendix} of \cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.

@@

We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Let $T_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
\begin{align*}
\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) + T_L(\hat{\lambda}_L) - G(0) \big).
\end{align*}
Lemma~\ref{lem:empiricprocforrho} shows that 
\[
\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L,
\]
where $X_L = \sqrt{L}\big( G_L(0) - T_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
\[
%\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1).
\]
It follows that
\[
%\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
\sqrt{L} m_L =  \rho_0 X_L + o_P(1). 
\]
Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$ where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.  Thus, the distribution of $\sqrt{L} m_L$ converges to the normal with zero mean and variance $\rho_0^2(p B_1 + d B_2)$ as required.  Because $X_L$ does not depend on $\hat{\lambda}_L$, it follows that $\covar(X_L, \sqrt{L}\hat{\lambda}_L) = 0$, and so,
\[
\covar(\sqrt{L}\hat{m}_L, \sqrt{L}\hat{\lambda}_L) = \covar(\rho_0 X_L + o_P(1), \sqrt{L}\hat{\lambda}_L) \rightarrow 0
\]
as $L \rightarrow \infty$.  It remains to prove the lemmas that we have used.  These lemmas are proved in Section~\ref{sec:proof-asympt-normappendix} of~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.

@@

Let the noise sequence $\{w_i\}$ is complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.  The joint density function of the real and imaginary parts is
\[
\frac{1}{2\pi\sigma^2}e^{-(x^2 + y^2)/\sigma^2}
\]
Theorem~\ref{thm:consistency}~and~\ref{thm:normality} hold and, since the distribution of $w_1$ is circularly symetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
%and the joint density function of the real and imaginary parts of $Z$ is
%\[
%p_Z(x,y) = \rho_0^2 p( \rho_0 (x - 1), \rho_0 y ) = \frac{\kappa^2}{2\pi}e^{-\kappa^2(x ^2 - 2x + 1 + y^2) }
%\] 
%where $\kappa = \tfrac{\rho_0}{\sigma}$.  
It can be shown that
\[
g(\phi) = \frac{\cos^2(\phi)}{4\sqrt{\pi}}\left( \frac{e^{-\kappa^2} }{\sqrt{\pi}} + \kappa \Psi(\phi)  e^{-\kappa^2\sin^2(\phi)}\cos(\phi) \right)
\]
where $\kappa = \rho_0/\sigma$ and
\[
\Psi(\phi) = 1 + \erf(\kappa \cos(\phi)) = 1 + \frac{2}{\sqrt{\pi}}\int_0^{\kappa \cos(\phi)} e^{-t^2} dt .
\]
The value of $A_1, A_2, B_1$ and $B_2$ can be efficiently computed by numerical integration using these formula.

@@

\begin{itemize}
\item Run simulations looking at what happens as the number of pilots changes.
\item Simulations showing what happens as $L$ changes for fixed number of pilots.
\item Show that sometimes just using pilots is better!
\end{itemize}

@@
